---
title: "ST-512-HW6-Yefrid_Cordoba"
format: html
editor: visual
author: "Yefrid Cordoba"
---

## Problem 1

Single factor random effects model, lager beer in a metropolitan area.

```{r}
beer_d <- read.table("HW6_Data.txt", header = TRUE,
                     colClasses = c("numeric", "factor", "factor")) 

```

```{r}
fit <- aov(sodium ~ brand, data = beer_d)
summary(fit)
```

### a.

For a random effects model:

$H_o : \sigma_\mu^2 = 0$\
$H_o : \sigma_\mu^2 \neq 0$\

Decision rule:\
We reject the null hypothesis if $P(F_{calc} > F_{critical})$.\

Assuming an $\alpha = 0.05$, and given that the probability of getting our $\sigma_\mu^2$ under the assumption of the null hypothesis is close to 0 (`P-value = <2e-6`), we reject the null hypothesis and conclude that the variability of sodium between brands beer brands is not zero $\sigma_\mu^2 \neq 0$.\

### b.

```{r}
tapply(beer_d$sodium, beer_d$brand, mean)
```
The values are:\
$y_{1\cdot} = 23.638$\
$y_{2\cdot} = 10.675$

### c.

```{r}
MSE <- 0.72
MStr <- 170.91
n <- 8
sm2 <- (MStr - MSE) / n
sm2
```

$E(\sigma^2) = MSE = 0.72$\
$E(\sigma_\mu^2) = \frac{MS_{Tr} - MSE}{n} = 21.274$

### d.

```{r}
sm2 / (MSE + sm2)
```

Estimated $\frac{\sigma_\mu^2}{\sigma^2+\sigma_\mu^2} = 0.967$

This value means that approximately 96.7% of the variation in sodium content of lager beers in a metropolitan area is due to their brand.

## Problem 2