---
title: "ST-512-HW6-Yefrid_Cordoba"
format: html
editor: visual
author: "Yefrid Cordoba"
---

## Problem 1

Single factor random effects model, lager beer in a metropolitan area.

```{r}
beer_d <- read.table("HW6_Data.txt", header = TRUE,
                     colClasses = c("numeric", "factor", "factor")) 

```

```{r}
fit <- aov(sodium ~ brand, data = beer_d)
summary(fit)
```

### a.

For a random effects model:

$H_o : \sigma_\mu^2 = 0$\
$H_o : \sigma_\mu^2 \neq 0$\

Decision rule:\
We reject the null hypothesis if $P(F_{calc} > F_{critical})$.\

Assuming an $\alpha = 0.05$, and given that the probability of getting our $\sigma_\mu^2$ under the assumption of the null hypothesis is close to 0 (`P-value = <2e-6`), we reject the null hypothesis and conclude that the variability of sodium between brands beer brands is not zero $\sigma_\mu^2 \neq 0$.\

### b.

```{r}
tapply(beer_d$sodium, beer_d$brand, mean)
```

The values are:\
$y_{1\cdot} = 23.638$\
$y_{2\cdot} = 10.675$

### c.

```{r}
MSE <- 0.72
MStr <- 170.91
n <- 8
sm2 <- (MStr - MSE) / n
sm2
```

$E(\sigma^2) = MSE = 0.72$\
$E(\sigma_\mu^2) = \frac{MS_{Tr} - MSE}{n} = 21.274$

### d.

```{r}
sm2 / (MSE + sm2)
```

Estimated $\frac{\sigma_\mu^2}{\sigma^2+\sigma_\mu^2} = 0.967$

This value means that approximately 96.7% of the variation in sodium content of lager beers in a metropolitan area is due to their brand.

## Problem 2

2-way random effects analysis of variance.

```{r}
a <- 8
b <- 5
n <- 4
SSA <- 1641
SSB <- 892
SSAB <- 1308
SSE <- 881
dfa <- a - 1
dfb <- b - 1
dfab <- dfa * dfb
dfe <- a * b * (n -1)

MSA <- SSA / dfa
MSB <- SSB / dfb
MSAB <- SSAB / dfab
MSE <- SSE / dfe

F_A <- MSA / MSAB
F_B <- MSB / MSAB
F_AB <- MSAB / MSE
```

| Source | sf | SS | MS | F-value |
|---------------|---------------|---------------|---------------|---------------|
| Rater (A) | [`r dfa`]{.underline} | 1641 | [`r round(MSA, 2)`]{.underline} | [`r round(F_A, 2)`]{.underline} |
| Product (B) | [`r dfb`]{.underline} | 892 | [`r round(MSB, 2)`]{.underline} | [`r round(F_B, 2)`]{.underline} |
| R\*P (AB) | [`r dfab`]{.underline} | 1308 | [`r round(MSAB, 2)`]{.underline} | [`r round(F_AB, 2)`]{.underline} |
| Error | [`r dfe`]{.underline} | 881 | [`r round(MSE, 2)`]{.underline} |  |

Using the formulas for the estimated mean squares for each of the sources of variations.

```{r}
sa <- (MSA - MSAB) / (b * n)
sa
sb <- (MSB - MSAB) / (a * n)
sb
sab <- (MSAB - MSE) / n
sab
s <- MSE
s
```

$\widehat{\sigma^2} = 7.34$\
$\widehat{\sigma_{\alpha\beta}^2} = 9.84$\
$\widehat{\sigma_{\alpha}^2} = 9.39$\
$\widehat{\sigma_{\beta}^2} = 5.51$

## Problem 3

