---
title: "ST-512-HW6-Yefrid_Cordoba"
format: html
editor: visual
author: "Yefrid Cordoba"
---

## Problem 1

Single factor random effects model, lager beer in a metropolitan area.

```{r}
beer_d <- read.table("HW6_Data.txt", header = TRUE,
                     colClasses = c("numeric", "factor", "factor")) 

```

```{r}
fit <- aov(sodium ~ brand, data = beer_d)
summary(fit)
```

### a.

For a random effects model:

$H_o : \sigma_\mu^2 = 0$\
$H_o : \sigma_\mu^2 \neq 0$\

Decision rule:\
We reject the null hypothesis if $P(F_{calc} > F_{critical})$.\

Assuming an $\alpha = 0.05$, and given that the probability of getting our $\sigma_\mu^2$ under the assumption of the null hypothesis is close to 0 (`P-value = <2e-6`), we reject the null hypothesis and conclude that the variability of sodium between brands beer brands is not zero $\sigma_\mu^2 \neq 0$.\

### b.

```{r}
tapply(beer_d$sodium, beer_d$brand, mean)
```

The values are:\
$y_{1\cdot} = 23.638$\
$y_{2\cdot} = 10.675$

### c.

```{r}
MSE <- 0.72
MStr <- 170.91
n <- 8
sm2 <- (MStr - MSE) / n
sm2
```

$E(\sigma^2) = MSE = 0.72$\
$E(\sigma_\mu^2) = \frac{MS_{Tr} - MSE}{n} = 21.274$

### d.

```{r}
sm2 / (MSE + sm2)
```

Estimated $\frac{\sigma_\mu^2}{\sigma^2+\sigma_\mu^2} = 0.967$

This value means that approximately 96.7% of the variation in sodium content of lager beers in a metropolitan area is due to their brand.

## Problem 2

2-way random effects analysis of variance.

```{r}
a <- 8
b <- 5
n <- 4
SSA <- 1641
SSB <- 892
SSAB <- 1308
SSE <- 881
dfa <- a - 1
dfb <- b - 1
dfab <- dfa * dfb
dfe <- a * b * (n -1)

MSA <- SSA / dfa
MSB <- SSB / dfb
MSAB <- SSAB / dfab
MSE <- SSE / dfe

F_A <- MSA / MSAB
F_B <- MSB / MSAB
F_AB <- MSAB / MSE
```

| Source | sf | SS | MS | F-value |
|:-------------:|:-------------:|:-------------:|:-------------:|:-------------:|
| **Rater (A)** | [`r dfa`]{.underline} | 1641 | [`r round(MSA, 2)`]{.underline} | [`r round(F_A, 2)`]{.underline} |
| **Product (B)** | [`r dfb`]{.underline} | 892 | [`r round(MSB, 2)`]{.underline} | [`r round(F_B, 2)`]{.underline} |
| **R\*P (AB)** | [`r dfab`]{.underline} | 1308 | [`r round(MSAB, 2)`]{.underline} | [`r round(F_AB, 2)`]{.underline} |
| **Error** | [`r dfe`]{.underline} | 881 | [`r round(MSE, 2)`]{.underline} |  |

Using the formulas for the estimated mean squares for each of the sources of variations.

```{r}
sa <- (MSA - MSAB) / (b * n)
sa
sb <- (MSB - MSAB) / (a * n)
sb
sab <- (MSAB - MSE) / n
sab
s <- MSE
s
```

$\widehat{\sigma^2} = 7.34$\
$\widehat{\sigma_{\alpha\beta}^2} = 9.84$\
$\widehat{\sigma_{\alpha}^2} = 9.39$\
$\widehat{\sigma_{\beta}^2} = 5.51$

## Problem 3

A mixed effects model where factor A (location) is the fixed effect, while factor B (player) is the random effect.

### a.

```{r}
SSA <- 1.30
SSB <- 3879.3
SSAB <- 323.67
SSE <- 15787.29
SST <- 19991.56

a <- 2
b <- 10
n <- 17

dfa <- a - 1
dfb <- b - 1
dfab <- dfa * dfb
dfe <- a * b * (n -1)
dfT <- a * b * n - 1

MSA <- SSA / dfa
MSB <- SSB / dfb
MSAB <- SSAB / dfab
MSE <- SSE / dfe

F_A <- MSA / MSAB
F_B <- MSB / MSE
F_AB <- MSAB / MSE

P_A <- pf(F_A, dfa, dfab, lower.tail = F)
P_B <- pf(F_B, dfb, dfe, lower.tail = F)
P_AB <- pf(F_AB, dfab, dfe, lower.tail = F)
```

| Source | df | SS | MS | F-value | P-value |
|------------|------------|------------|------------|------------|------------|
| **Location (A fixed)** | [`r dfa`]{.underline} | 1.30 | [`r round(MSA, 2)`]{.underline} | [`r round(F_A, 2)`]{.underline} | `r P_A` |
| **Player (B random)** | [`r dfb`]{.underline} | 3879.30 | [`r round(MSB, 2)`]{.underline} | [`r round(F_B, 2)`]{.underline} | `r P_B` |
| **Player \* Location** | [`r dfab`]{.underline} | 323.67 | [`r round(MSAB, 2)`]{.underline} | [`r round(F_AB, 2)`]{.underline} | `r P_AB` |
| **Error** | [`r dfe`]{.underline} | 15787.29 | [`r round(MSE, 2)`]{.underline} |  |  |
| **Total** | [`r dfT`]{.underline} | 19991.56 |  |  |  |

### b.

Testing:

$H_o : \sigma_{\alpha\beta}^2 = 0$\
$H_a : \sigma_{\alpha\beta}^2 \neq 0$\

Given the P-value for the interaction term (0.6823), and assuming an $\alpha = 0.05$, the probability is higher that our threshold. Therefore we fail to reject the null hypothesis and conclude that we don't have enough evidence to suggest that there is an interaction between player and location.

### c. 

Testing:

$H_o : \alpha_1 = \alpha_2 = 0$\
$H_a : at~least~one~\alpha_i \neq 0$\

Given the P-value for the location term (0.8534), and assuming an $\alpha = 0.05$, the probability is higher that our threshold. Therefore we fail to reject the null hypothesis and conclude that we don't have enough evidence to suggest that location has a main effect on the mean points gotten by a player.

### d.

Testing:

$H_o : \sigma_{\beta}^2 = 0$\
$H_a : \sigma_{\beta}^2 \neq 0$\

Given the P-value for the player effect (9.424e-12), and assuming an $\alpha = 0.05$, the probability is lower than our threshold. Therefore we reject the null hypothesis and conclude that we have enough evidence to suggest that the player has a main effect on the average points.

### e.

```{r}
sab <- (MSAB - MSE) / n
sab
sb <- (MSB - MSE) / (a * n)
sb
s <- MSE
s
```

$\widehat{\sigma^2} = 49.335$\
$\widehat{\sigma_{\alpha\beta}^2} = 0$\
As statistically proven, the $\widehat{\sigma_{\alpha\beta}^2}$ is not estimated to be different than 0, since a variance cannot be negative, its value is set to 0.\
$\widehat{\sigma_{\beta}^2} = 11.226$\

## Problem 4

